%!TEX root = Economics and Behaviour.tex


\section{Extensive form}

From a game-theoretic viewpoint, the strategic and the extensive form are merely two different representations of the same strategic decision situation. But while the strategic form often only provides a static description of a game, the extensive form provides additional properties like move order or level of information in a game.

\begin{example}[Dictator-Game] \label{Dictator-Game} \index{Dictator-Game}
Imagine two players, a proposer $P$ can split up 10 \euro ~ (up to \euro-level) between him and a receiver $R$. 
	\begin{itemize}
		\item Question 1: Assume for a minute $P$ is totally selfish and only cares about his own profits. Is there a strictly dominant strategy for $P$? \\
			Yes! $(10, 0)$ (money proposer, money receiver) is strictly dominant.
		\item Question 2: What if $P$ is a pure altruist and just cares about the money $R$ receives? \\
			Then $(0, 10)$ is strictly dominant.
	\end{itemize}
\end{example}

This simply example for a game we are now going to extend into a dynamic game with the addition that after the proposal the receiver can depending on the offer either accept or reject the proposal.

\begin{definition}[extensive form]
	A sequential game like the \hyperref[Dictator-Game]{Dictator-Game} is suitably modelled by an \begriff{extensive form} and we will later find out why it commonly has the shape of a tree. The definition of a game in extensive form consists of:
\begin{itemize}  \index{node}
	\item A finite set of nodes $\mathcal{X}$, a finite set of possible actions $\mathcal{A}$ and a finite set of players ${1, \dotsc, N}$
	\item A function $p \colon \mathcal{X} \rightarrow \{ \mathcal{X} \cup \emptyset \}$ specifying a single immediate predecessor of each node $x \in \mathcal{X}$; except for $x_{0}$ which is the \begriff{initial node}. The immediate successor nodes of $x$ are $s(x) = p^{-1}(x)$. \\
		To have a tree structure, a predecessor of x can never be a successor of x and vice versa. The set of \begriff{terminal nodes} is $T = \{ x \in \mathcal{X}: s(x) = \emptyset \}$. All other nodes $\mathcal{X} \setminus T$ are \begriff{decision nodes}.
	\item A function $\alpha \colon \mathcal{X} \setminus \{ x_{0} \} \rightarrow \mathcal{A}$ giving the action that leads to any non-initial node $x$ from its immediate predecessor $p(x)$ with
		\[ x', x'' \in s(x); x' \neq x'' \Rightarrow \alpha(x') \neq \alpha(x''). \]
		The set of choices at decision node $x$ is $c(x) = \{ a \in \mathcal{A} : a = \alpha(x')$ for some $x' \in s(x) \}$
	\item A collection of information sets $\mathcal{H}$, and a function $H : \mathcal{X} \rightarrow \mathcal{H}$, assigning each decision node $x$ to an information set $H(x) \in \mathcal{H}$ with $c(x) = c(x')$ if $H(x) = H(x')$.
		The choices available at information set $H$ can be written as		\[ C(H) = \{a \in A : a \in c(x) \text{ for } x \in H \}. \]
	\item A function $\tau : \mathcal{H} \rightarrow \{0, 1, \dotsc, N \}$ assigning a player to each information set ($i = 0$ 'nature').
		The collection of player $i$'s information set is denoted by $\mathcal{H}_i = \{H \in \mathcal{H} : i = \tau(H) \}$.
	\item A function $\rho : \mathcal{H}_{0} \times \mathcal{A} \rightarrow [0, 1]$ assigning a probability to each action of nature with $ \rho (H,a) = 0$ if $a \notin C(H)$ and $\sum_{a \in C(H)} \rho(H,a) = 1$ for all $H \in \mathcal{H}_{0}$.
	\item A collection of payoff functions $u = \{ u_{1}(\dotsc), \dotsc, u_{N} (\dotsc) \}$, where $u_{i} \colon T \rightarrow \mathds{R}$.
\end{itemize}
\end{definition}

Now, back to our example:

\begin{example}[Ultimatum-Game]	\label{ultimatumgame} \index{Ultimatum-Game} 
Given two players, one of them, the proposer $P$, receives 10 \euro~and decides how to divide the sum between himself and another player, the receiver $R$;  where $P$ receives $x_{p}$ (with $x_{p} \in \{ 0, 1, 2, \dotsc, 10 \}$) and $R$ collects the remaining $10 - x_{p}$. After the proposer's decision the responder can either accept or reject this offer, if he accepts the money is split according to the proposal, if he rejects neither player receives any money.

To sum it up:
	\begin{itemize}
		\item $P$ proposes split up $(x_{p}, 10 - x_{p})$
		\item $R$ accepts or rejects
			\begin{itemize}
				\item If $R$ accepts $(a)$ proposal becomes implemented, $P$ receives $x_{p}$ and R $10 - x_{p}$
				\item If $R$ rejects $(r)$ the payout for both players is $0$.
			\end{itemize}
	\end{itemize}
	
	\begin{center}
		\begin{tikzpicture}[auto, level distance=30mm, sibling distance=35mm]
			\node [circle,draw] (z){$P$}
 						child {node [circle,draw] (a) {$10$}	}
 						child {node [] (b) {$ $}	}
 						child {node [circle,draw] (c) {$x_{p}$}
  							child {node [circle,draw] (f) {$(x_{p}, 10 - x_{p})$}	}
  			 				child {node [circle,draw] (g) {$(0, 0)$}	}			
 						}
 						child {node [] (d) {$$}	}
  						child {node [circle,draw] (e) {$0$}
			};
			\path (a) -- (b) node [midway] {$\dotsc$};
			\path (b) -- (c) node [midway] {$\dotsc$};
			\path (c) -- (d) node [midway] {$\dotsc$};
			\path (d) -- (e) node [midway] {$\dotsc$};
			\path (f) -- (g) node [midway] {$a$ $/$ $r$};
		\end{tikzpicture}				
	\end{center}
	
	A strategy set in this game, as it needs to specify a complete plan of actions, should be similar to
	\begin{itemize}
		\item Proposer sets an $x_{p}$
		\item Receiver decides for \textit{any} $x_{p}$ that might come up as an offer if he'd accept or reject it.
	\end{itemize}
	
	\textbf{1. Question:} Can the outcome $(5, 5)$ be stabilised as a Nash-Equilibrium? \\
	\textbf{Answer:} Yes. Let's say P proposes $x_{p} = 5$ and the strategy set for R is defined by accepting any value of $x_{p} \leq 5$ and rejecting the offer for values larger than $5$. \\
		In this situation $(5, 5)$ would state a Nash-Equilibrium.


	\textbf{2. Question:} Is there another Nash-Equilibrium that stabilises the $(5, 5)$ outcome? \\
	\textbf{Answer:} Yes. If P proposes $x_{p} = 5$ and R now accepts the offer only for $x_{p} = 5$, he'd therefore reject any other case ($x_{p} \neq 5$). 
	
	
	\textbf{3. Question:} Can $(0, 10)$ be stabilised as a Nash-Equilibrium? \\
	\textbf{Answer:} Yes. We set for P the proposal to $x_{p} = 0$ and define R's strategy set as accepting for $x_{p} = 0$, and rejecting for any other case ($x_{p} \geq 1$). 
\end{example}
As we can see the Nash-Equilibrium can lead to an infinite amount of outcomes some of them even with implausible threats. As this seems unsatisfying we'd therefore like to refine once again our set of equilibria which leads us to the (subgame) perfect Nash-Equilibrium:

\begin{definition}[Subgame] \index{subgame}
	A subgame is any part of a game that meets the following criteria:
	\begin{itemize}
		\item It has a single initial node that is the only member of that node's information set (i.e. the initial node is in a singleton information set).
		\item If a node is contained in the subgame then so are all of its successors
		\item If a node in a particular information set is in the subgame then all members of that information set belong to the subgame.
		\item and finally the node must not contain a deterministic state but instead at least one non-trivial choice
	\end{itemize}
\end{definition}

\begin{definition}[(Subgame) Perfect Nash-Equilibrium] \index{subgame perfect Nash-Equilibrium} 
A strategy profile is a subgame perfect Nash-Equilibrium if it represents a Nash equilibrium of every subgame of the original game. Informally, this means that if the players played any smaller game that consisted of only one part of the larger game and their behaviour represents a Nash equilibrium of that smaller game, then their behaviour is a subgame perfect equilibrium of the larger game. 
\end{definition}
How to find a Perfect Nash-Equilibrium:
\begin{enumerate}
	\item Define (one set of) optimal actions for the last subgame
	\item Replace that decision nodes with the respective outcome
	\item Repeat $(1)$ and $(2)$ until the initial decision node is reached.
\end{enumerate}

\begin{example}[Sequel to the Ultimatum-Game]
	Searching for the Perfect Nash-Equilibrium in this case leads to:
	\begin{enumerate}
		\item Defining the optimal actions for the responder
			\begin{itemize}
				\item Given the case of $P$ offering $0$ ($x_{p} = 10$), $R$ receives nothing and is therefore indifferent between refusing and accepting. Let's assume for now he'd accept.
				\item For $x_{p} \in [0, 10)$ $R$ always has to accept the offer since $10 - x_{p} > 0$.
			\end{itemize}
		\item Assuming $R$ always accepts, we now can reduce the game
			\begin{center}
				\begin{tikzpicture}[level/.style={sibling distance=25mm/#1}, level 	distance=25mm]
					\node [circle,draw] (z){$P$}
 						child {node [circle,draw] (a) {$(10, 0)$}	}
 						child {node [circle,draw] (b) {$(9, 1)$}	}
 						child {node [] (c) {$ $}	}
 						child {node [circle,draw] (d) {$(5, 5)$}	}
 						child {node [] (e) {$ $}	}
  						child {node [circle,draw] (f) {$(0, 10)$}
					};
					\path (b) -- (c) node [midway] {$\dotsc$};
					\path (c) -- (d) node [midway] {$\dotsc$};
					\path (d) -- (e) node [midway] {$\dotsc$};
					\path (e) -- (f) node [midway] {$\dotsc$};
				\end{tikzpicture}				
			\end{center}
		\item A simple search in this reduced game tree for the strategy with the highest expected payoff for $P$ yields the subgame perfect Nash-Equilibrium as we have already reached the initial node.
	\end{enumerate}
	$\Rightarrow P$ playing $x_{p} = 10$ together with $R$ always accepting constitutes a subgame perfect Nash-Equilibrium. 

	Now changing our initial assumption about $R$'s decision in case of an $x_{p} = 0$ offer causes (even if the game tree changes only slightly) a different subgame perfect Nash-Equilibrium
			\begin{itemize}
				\item Again, for $x_{p} \in [0, 10)$ $R$ receives $10 - x_{p} > 0$, therefore he'd have to accept the offer in these situations.
				\item For the remaining offer $0$ ($x_{p} = 10$), it would be just as plausible to refuse the offer, since he is indifferent between both choices. Hence, accepting for $x_{p} \in [0, 10)$ and refusing for $x_{p} = 10$ is also a subgame perfect strategy.
			\end{itemize}
		 Our reduced game tree looks now like this:
			\begin{center}
				\begin{tikzpicture}[level/.style={sibling distance=25mm/#1}, level 	distance=25mm]
					\node [circle,draw] (z){$P$}
 						child {node [circle,draw] (a) {$(0, 0)$}	}
 						child {node [circle,draw] (b) {$(9, 1)$}	}
 						child {node [] (c) {$ $}	}
 						child {node [circle,draw] (d) {$(5, 5)$}	}
 						child {node [] (e) {$ $}	}
  						child {node [circle,draw] (f) {$(0, 10)$}
					};
					\path (b) -- (c) node [midway] {$\dotsc$};
					\path (c) -- (d) node [midway] {$\dotsc$};
					\path (d) -- (e) node [midway] {$\dotsc$};
					\path (e) -- (f) node [midway] {$\dotsc$};
				\end{tikzpicture}				
			\end{center}
	$\Rightarrow P$ playing $x_{p} = 9$ together with $R$ always accepting if $x_{p} < 10$ and refusing for $x_{p} = 10$ also constitutes a subgame perfect Nash-Equilibrium. 
\end{example}



\newpage